Bayesian exponential random graph modeling of KWL Burial network

### 2020-12-11 analysis/paper/paper.Rmd
edit discussion and add references

<!--fit in a big anthropological picture in a periclonial context -->
The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017]. Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007]. Despite relatively weak control, northeastern Taiwan was incorporated into colonial territories in 1647 after the Dutch conquests []. The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies [@Kang2016]. At Kiwulan, after the arrival of Europeans, some individuals may be differentiated or privileged to possess more distinctive goods through becoming the trade partners with the Europeans. Changes in interburial networks we observed may result from differential access to exotic trade relative to the European colonization and gradually led to an increased social differentiation. The image of European power could be embedded into prestige objects pre-existing in Indigenous culture where their values were amplified with concepts of wealth or power, since good values are contextually mutable and entangled in historical processes [@Aswani2003; @Thomas2009]. This may lead to ambitious individuals to compete with each other and accumulate trade goods. It demonstrates the agency of local Indigenous societies to incorporate and manipulate trade items in a precolonial context, gradually resulting in social changes from a more corporate to a more network strategy.

>>>>

<!--fit in a bigger historical picture in a periclonial context -->
The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017]. Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007]. In 1626, the Spanish built their forts in Keelung and Tamsui, northern Taiwan, and later defeated and took over by the Dutch in 1642 [@Andrade2007]. Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attached by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016]. The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.

<!--anthopological implication at Kiwulan and limitation-->
At Kiwulan, after the European presence, some individuals may be differentiated or privileged to possess more distinctive goods through becoming the trade partners with the Europeans. Changes in interburial networks we observed may result from differential access to exotic trade relative to the European colonization and gradually led to an increased social differentiation. The image of European power could be embedded into prestige objects pre-existing in Indigenous culture where their values were amplified with concepts of wealth or power, since good values are contextually mutable and entangled in historical processes [@Aswani2003; @Thomas2009]. This may lead to ambitious individuals to compete with each other for the goods and accumulate wealth. It also demonstrates the agency of local Indigenous societies to incorporate and manipulate trade items in a precolonial context with weak colonial control, gradually resulting in social changes from a more corporate to a more network strategy. Our interpretation is based on relations indicated by burial data that can be supported by the clustered distribution of trade goods in Kiwulan residential area [eg. @LWandBM2020ornament]. We would still need more other different evidence to understand the process of social change. In addition, we acknowledge the fragmentation issue in archaeology data that might limit network interpretation [eg. @Erik2015], such as missing age and sex nodal attributes for some burials in our data. Despite we lack of informative results about their effect on network formation, the Bayesian approach still provides effective and reliable statistical inferences on the overall network properties by reducing the missing data effect.
